---
title: "Class Reference"
author: "Student name"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.


## Things I Learned on J

The command to set a working directory is setwd(). For example:

```{r}
setwd("~/Documents/GitHub/data_journalism_2022_spring")
```

```{r}
install.packages("tidyverse")
library(tidyverse)
```


### Summarizing

I need to use group_by and summariz. Here's an example of grouping by county and calculating counts, sum and other descriptive statistics.


```{r}
ppp_maryland_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n(),
    total_loans_amount = sum(amount),
    mean_loan_amount = mean(amount),
    median_loan_amount = median(amount),
    min_loan_amount = min(amount),
    max_loan_amount = max(amount)
  ) %>%
  arrange(desc(max_loan_amount))
```


### Interesting Packages

~ library(tidyverse) - The tidyverse is a collection of open source packages for the R programming language introduced by Hadley Wickham and his team that "share an underlying design philosophy, grammar, and data structures" of tidy data

~ library(lubridate) - Lubridate is an R package that makes it easier to work with dates and times. 

### September 19, 2022

### Commands

my_new_variable -> [all the code that summarizes the data]

[name] <- read_csv - Reads CSV file

head(name) - shows headings

summary(name) - summarizes data

colnames(name) - shows column names

glimpse(name) - makes a tibble with all the sorted information

select (name, different name, etc) - only shows these columns

mutate(new column being created = column it was based on) - add a new column based on an existing column; based on a calculation using two existing columns
   - mutate is not permanent unless you save it to a new variable