---
title: "Data Analysis"
author: "Abby Wallace, Lisa Woelfl, Torrence Banks, Noah Ferguson"
date: "11/11/2022"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are analyzing data from the Maryland Lobbying Registrations. [https://lobby-ethics.maryland.gov/]

## Load libraries
Loading required libraries for this analysis.
```{r echo=FALSE, message=FALSE}
# Load the tidyverse here
library(tidyverse)
# Load janitor here
library(janitor)
library(lubridate)
library(stringr)
library(dplyr)
library(tidyr)
library(ggthemes)
```

## Load and Cleaning Data
Load required data
Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")
```{r}
# Loading the dataframes and binding them together
one <- read_csv("data/registrations_dataset/registrations(1).csv") %>% clean_names()
two <- read_csv("data/registrations_dataset/registrations(2).csv") %>% clean_names()
three <- read_csv("data/registrations_dataset/registrations(3).csv") %>% clean_names()
four<- read_csv("data/registrations_dataset/registrations(4).csv") %>% clean_names()
five <- read_csv("data/registrations_dataset/registrations(5).csv")%>% clean_names()
six <- read_csv("data/registrations_dataset/registrations(6).csv")%>% clean_names()
seven <- read_csv("data/registrations_dataset/registrations(7).csv")%>% clean_names()
eight <- read_csv("data/registrations_dataset/registrations(8).csv")%>% clean_names()
nine <- read_csv("data/registrations_dataset/registrations(9).csv")%>% clean_names()
total_registrations <- rbind(one, two, three, four, five, six, seven, eight, nine)
write_csv(total_registrations, "data/registrations_dataset/total_registrations.csv")
# We cleaned the data in OpenRefine and split the date column
clean_total_registrations <- read_csv("data/registrations_dataset/clean_total_registrations.csv")
clean_total_registrations[c('start', 'end')] <- str_split_fixed(clean_total_registrations$registration_period, '-', 2)
clean_total_registrations <- clean_total_registrations %>%
  mutate(start = mdy(start)) %>%
  mutate(end = mdy(end))
```

## Basic explorations

```{r}
glimpse(clean_total_registrations)
```
Rows: 26,580
Columns: 10

From accessing the Maryland Lobbying Registrations website, we downloaded 9 CSVs of registration data points. We then cleaned each individually and binded them together to create a comprehensive cleaned CSV and data set. We added cleaned columns for organization name & employer and split the date column to have a start and end date for the registration. Our comprehensive, cleaned data set includes 26,580 rows and 10 columns including form_id for a registration number, lobbyist_registrant for the name of the lobbyist, clean_lobbyist_registrant for a standardized list of the lobbyists by name, organization_firm for the name of the organization the lobbyist works for, clean_organization_firm for a standardized list of the organizations the lobbyists work for, employer for what company hired the lobbyist, clean_employer for a standardized list of companies that hired their corresponding lobbyist, registration_period for the registration period that the lobbyist was registered to lobby on behalf of the company listed, start for the start date of the registration period and end for the end of the registration period. Our earliest listed start date is January 1, 2016 and our most recent listed end date is December 31, 2021.

After we merged the data sets into total_registrations, we exported a CSV file of the information and cleaned the data in OpenRefine. There only remained a handful of names that were formatted slightly differently that we needed to clean before exporting the now cleaned CSV back to RStudio where we created the cleaned name columns and the individual start and end date columns.

Limitations: The dataframe is pretty basic with a small number of columns. That means we will have to cross-reference the Maryland Lobbying Registrations and other external websites if we want to know how much a lobbyist is making or how much an employer is spending on lobbying. The standard dataset doesn't even include the registration matters. Another limitation is that the earliest data is from 2016, which means we can't look at any trends that surpass 6 years.

On the Maryland Lobbying Registrations website, they provide a search tool titled "Registration Matters" where you can search registrations by the category/topic. For our special topics, we're downloading the CSV files from their pre-defined "Energy", "Gaming" and "Criminal Law Procedure" categories online. The datasets for "Registration Matter" only goes back to 2019.

To answer most of our questions, we will have to do additional research. For example, we can ask our data for the number of employers and lobbyists in the energy sector and see how that changed over the last few years. To examine how much the top firms make, we have to go back to Maryland Lobbying Registrations and potentially research beyond just the single website.

### Question 1

**Question text**: Who earned the most money in the latest complete lobbying year (Nov 21 - Oct 22) and which employer spent the most?

In the MSEC activity reports, lobbyists have to report the compensation they have received. The employers state their total expenses.

Limitations: While the data for lobbyist registrations can be filtered by registration matter, the same isn't possible for the activity reports.

**Analysis summary**

Matthew Celentano is by far the top earner for the latest complete lobbying year (Nov 21 - Oct 22). He was compensated over 100 million dollars by the League of Life & Health Insurers of Maryland, Inc. Celentano works for the firm Funk & Bolton, P.A.. According to his LinkedIn profile, he is also the executive director of League of Life, which is a trade association for the insurance industry. (Unsurprisingly, League of Life is the employer that spent the most money in this lobbying year. It's the only employer spending over a million. Second place goes to Constellation Energy Generation, LLC, which spent over $800,000.)

In the years before, Celentano made much less money, between $190,000 and $300,000. We couldn't find any explanations for the sharp rise in the spending of League of Life, which has been paying Celentano for lobbying at least since 2019. There aren't any relevant news stories about League of Life or Celentano. The extremely high amount spent could indicate a simple typo or a bigger story. Either way, this finding is worth investigating.

Overall, 12 lobbyists made over a million, while 174 made less than one thousand dollars, 131 of those were compensated $0.

Maryland Veterinary Medical Association	spent the least money of all employers with just over $18,000.

```{r}
## We fix the date
activity_reports <- read_csv("data/activity_reports_dataset/activity_reports_nov_21_oct_22.csv") %>% clean_names()
activity_reports[c('start', 'end')] <- str_split_fixed(activity_reports$reporting_period, '-', 2)
activity_reports <- activity_reports %>% 
  mutate(new_start = as.Date(start, format = "%m/%d/%y")) %>% 
  mutate(new_end = as.Date(end, format = "%m/%d/%y")) %>% 
  mutate(start_year = floor_date(new_start, "year"))
activity_reports

## And transform the relevant columns to numeric
clean_activity_reports <- activity_reports %>% 
  mutate(total_expenses = gsub(",","",total_expenses)) %>% 
  mutate(total_expenses = str_remove(total_expenses, "[$]")) %>% 
  mutate(total_expenses = as.numeric(total_expenses)) %>% 
  mutate(lobbyist_compensation = gsub(",","",lobbyist_compensation)) %>% 
  mutate(lobbyist_compensation = str_remove(lobbyist_compensation, "[$]")) %>% 
  mutate(lobbyist_compensation = as.numeric(lobbyist_compensation))
clean_activity_reports

## Now we want to see the top earners among the lobbyists
top_lobbyists <- clean_activity_reports %>% 
  group_by(lobbyist_registrant) %>% 
  summarise(lobbyist_compensation = sum(lobbyist_compensation)) %>% 
  arrange(desc(lobbyist_compensation))
top_lobbyists
over_million <- top_lobbyists %>% 
  filter(lobbyist_compensation > 1000000)
over_million
less_thousand <- top_lobbyists %>% 
  filter(lobbyist_compensation < 1000)
less_thousand
zero <- top_lobbyists %>% 
  filter(lobbyist_compensation == 0)
zero
celentano <- clean_activity_reports %>% 
  filter(lobbyist_registrant == "Celentano, Matthew")
celentano

## Now we want to know which employers spend the most money
top_employers <- clean_activity_reports %>% 
  group_by(employer) %>% 
  summarise(total_expenses = sum(total_expenses)) %>% 
  arrange(desc(total_expenses))
top_employers

## Let's take a closer look at Celentano's activity reports
celentano_reports <- read_csv("data/activity_reports_dataset/celentano_activity_reports.csv") %>% clean_names()
celentano_reports[c('start', 'end')] <- str_split_fixed(celentano_reports$reporting_period, '-', 2)

## We fix the dates
celentano_reports <- celentano_reports %>% 
  mutate(new_start = as.Date(start, format = "%m/%d/%y")) %>% 
  mutate(new_end = as.Date(end, format = "%m/%d/%y")) %>% 
  mutate(start_year = floor_date(new_start, "year"))
celentano_reports

## And convert the relevant columns to numeric
clean_celentano_reports <- celentano_reports %>% 
  mutate(total_expenses = gsub(",","",total_expenses)) %>% 
  mutate(total_expenses = str_remove(total_expenses, "[$]")) %>% 
  mutate(total_expenses = as.numeric(total_expenses)) %>% 
  mutate(lobbyist_compensation = gsub(",","",lobbyist_compensation)) %>% 
  mutate(lobbyist_compensation = str_remove(lobbyist_compensation, "[$]")) %>% 
  mutate(lobbyist_compensation = as.numeric(lobbyist_compensation))
clean_celentano_reports

## We want to see how much Celentano reported each year
clean_celentano_reports %>% 
  group_by(start_year) %>% 
  summarise(sum(lobbyist_compensation))

## When did he start working for League of Life?
clean_celentano_reports %>% 
  filter(employer == "League of Life & Health Insurers of Maryland, Inc.")
```


### Question 2
* **Question text**: Who are the top three "gaming" lobbyists with the most clients (employers), as described and categorized by Maryland Lobbying Registrations? How many of the gaming lobbyists are still active after sports betting was approved by voters November 3, 2020? Who are the top three organizations employing lobbyists and do any of them have locations already within Maryland?

Limitations: Online sports betting was only legalized and began November 23, 2022, so there is not a lot of time to consider that element. Voters approved legal sports betting in Maryland in November 2020, and state legislators passed bills facilitating wagering in April 2021. In-person legalized sports betting did not actively start until December 2021. The earliest registration date is November 2019, so we only have about three-to-four years of data.

* **Analysis summary**: I first downloaded the data set provided by Maryland Lobbying Registrations when you set the "Registration Matters" element of the internal search engine to "Gaming." Upon downloading, there were 328 registrations, and I cleaned the names of lobbyists, organization firms and employers using OpenRefine. After downloading the cleaned data set, I uploaded the set to RStudio. 

Ivan V. Lanier, David H. Carroll, Jr. and Sushant Sidh are the top three lobbyists in Maryland with the most lobbying clients since 2019 with Lanier having 26 registration contracts and Carroll and Sidh both having 13 registration contracts each. Lanier actually founded Greenwill Consulting Group, LLC and operates as the President and Chief Executive Officer where he specializes in lobbying for energy, medical cannabis, transportation, telecommunications, property and health-care. Carroll is a founder and principal lobbyist for Capitol Strategies, LLC and has a history lobbying on issues such as transportation, energy, gaming, education, health care, tax issues, environmental regulation, telecommunications and procurement. Sidh is also a principal lobbyist for Capitol Strategies, LLC and used to work as a Deputy Chief of Staff to Maryland Governor Parris N. Glendening and spent six years in the administration. 

There are 261 registration period contracts regarding gaming that were/are active now that sports betting is legal in Maryland from 49 different gaming employers. Sports betting was legalized in November 2020, began in-person in December 2021 and online sports-betting began November 2022. Most of these contracts are approximately one year long. As for gaming employers that have had the most lobbyist contracts since 2019, CBAC Gaming, LLC (Horseshoe Casino) is at the top with 36 registration period contracts with a tie for second place between MGM Resorts International and PPE Casino Resorts Maryland, LLC with 20 each. CBAC Gaming, LLC owns the Horseshoe Casino in Baltimore which is two stories and the second largest casino in Maryland. The Horseshoe Baltimore Casino features a Caesars Sportsbook on the first place which opened in December 2021 and offers all-year round sports betting. MGM Resorts International has a location in National Harbor in Oxon Hill and features a luxury resort, retail, dining, entertainment and casino. MGM National Harbor also has BetMGM Sportsbook & Lounge for sports betting. PPE Casino Resorts Maryland, LLC owns Live! Casino & Hotel Maryland in Hanover which is Maryland's largest casino and boasts a FanDuel Sportsbook center for sports betting.

Since CBAC Gaming, LLC (Horseshoe Casino) has the most lobbyists and active contracts since 2019, we analyzed how much money it spent per year. Since most of the lobbying periods start around October, 2022 has the lowest amount spent so far as it is an incomplete year. CBAC Gaming spent an equal amount of money in 2020 and 2021 of $240000 per year. Since in-person sports betting was initially passed in 2020, and online sports betting was passed in 2022 that could explain the uptick in lobbying spending.

```{r}
### I first uploaded the dataset and cleaned the names up, split the registration periods into start and end dates, and changed them to be categorized as dates rather than as characters. Now, I can work with this data set.
gaming <- read_csv("data/gaming.csv") %>% clean_names()
gaming[c('start', 'end')] <- str_split_fixed(gaming$registration_period, '-', 2)
gaming_dates <- gaming %>%
  mutate(start = mdy(start)) %>%
  mutate(end = mdy(end))
gaming_dates
```

```{r}
#In order to determine which lobbyists have the most clients, I first group by the lobbyists themselves and count how many registration period contracts they each have within the data set. Here, I can see that Ivan V. Lanier, David H. Carroll, Jr. and Sushant Sidh are the top three lobbyists in Maryland with Lanier having 26 registration contracts and Carroll and Sidh both having 13 registration contracts each.
top_gaming <- gaming_dates %>%
  group_by(lobbyist_registrant, organization_firm) %>%
  count() %>%
  arrange(desc(n))
top_gaming
```
```{r}
# Taking a closer look at Lanier, we can filter the dataset to only include his registrations and count how many registration contracts he has with each organization that employed him. We can see that since 2019, Lanier has worked with 11 different employers as a lobbyist for them with Golden Entertainment and HPSO Investments LLC, T/A Riverboat on the Potomac tied for Lanier's top client with 4 contracts each.
lanier_gaming <- gaming_dates %>% filter(lobbyist_registrant == "Lanier, Ivan  V.") %>%
  count(employer, source = TRUE) %>%
  arrange(desc(n))
lanier_gaming
```

```{r}
# As for Carroll, we can repeat the process above and see that Carroll has only had four different organizations hire him to be a lobbyist since 2019. Carroll's top employer for lobbying is tied between Atlantic Bingo Supply, Inc and Delta Bingo t/a Daily Double Bingo with four contracts each.
carroll_gaming <- gaming_dates %>% filter(lobbyist_registrant == "Carroll, Jr., David H.") %>%
  count(employer, source = TRUE) %>%
  arrange(desc(n))
carroll_gaming
```
```{r}
# For our last lobbyist, we can see that Sidh and Carroll have the exact same history regarding frequency of lobbying and agencies. When we look at our count of lobbyists, their organizations and how many registration periods they have been active for, we can see that Carroll and Sidh work for the same lobbying agency.
sidh_gaming <- gaming_dates %>% filter(lobbyist_registrant == "Sidh, Sushant") %>%
  count(employer, source = TRUE) %>%
  arrange(desc(n))
sidh_gaming
```
```{r}
#As for determining how many contracts are still active, I set the "end" filter for the day sports betting became legal in Maryland on November 03, 2020. We can see that there are 261 registration period contracts regarding gaming that are still active now that sports betting is legal in Maryland.
gaming_contracts <- gaming_dates %>%
  arrange(desc(end)) %>%
  filter(end >= "2020-11-03")
gaming_contracts
```
```` {r}
# Of these 261 active contracts, we can see here that they are from 49 different employers with CBAC Gaming, LLC (Horseshoe Casino) having the most active contracts at 27.
gaming_contracts %>%
  count(employer, source = TRUE) %>%
  arrange(desc(n))
```
```{r}
#Taking a step back, we can see how many registration period contracts each gaming employer had since 2019. CBAC Gaming, LLC (Horseshoe Casino) clearly is at the top with 36 registration period contracts with a tie for second place between MGM Resorts International and PPE Casino Resorts Maryland, LLC with 20 each.
employer_hires <- gaming_dates %>%
  group_by(employer) %>%
  count() %>%
  arrange(desc(n))
employer_hires
```
```{r}
#To visualize this data, we can use ggplot to illustrate just how many more lobbyists/registration period contracts CBAC Gaming, LLC (Horseshoe Casino) truly had.
top_hires <- employer_hires %>%
  head(10)
top_hires %>%
  ggplot() +
   geom_bar(aes(x=reorder(employer,n), weight=n)) +
   coord_flip() + 
   theme_economist() +
   labs(
     title="Number of Lobbying Periods",
     y = "Number of Lobbying Registration Periods",
     x = "Organization Employing Lobbyists",
     caption = "Maryland Lobbying Registrations")
```
```{r}
# Taking a closer look at CBAC Gaming, LLC, we can search through their activity reports by searching that employer/company on Maryland Lobbying Registrations. First, I downloaded the dataset and cleaned the data. I split the dates into two independent columns.
cbac_activity_reports <- read_csv("data/activity_reports_dataset/cbac_activity_reports.csv") %>% clean_names()
cbac_activity_reports[c('start', 'end')] <- str_split_fixed(cbac_activity_reports$reporting_period, '-', 2)
```
```{r}
#Next, I coded the start and end columns to categorize the dates as such rather than as characters.
cbac_activity_reports <- cbac_activity_reports %>%
  mutate(start = mdy(start)) %>%
  mutate(end = mdy(end))
```

```{r}
# After mutating the relevant lobbyist_compensation columns, the columns became money rather than characters. We were then able to create a floor_date for each year and observe the amount of money spent by CBAC Gaming, LLC each year since 2019.
cbac_activity_years <- cbac_activity_reports %>%
  mutate(lobbyist_compensation = gsub(",","",lobbyist_compensation)) %>%
  mutate(lobbyist_compensation = str_remove(lobbyist_compensation, "[$]")) %>%
  mutate(lobbyist_compensation = as.numeric(lobbyist_compensation)) %>%
  mutate(start_year = floor_date(start, "year")) %>%
  group_by(start_year) %>%
  summarise(total = (sum(lobbyist_compensation)))
cbac_activity_years
```

```{r}
# To visualize the data, I created another bar graph and organized it by the highest amount spent per year since 2019.
cbac_activity_years %>%
  ggplot() +
   geom_bar(aes(x=reorder(start_year,(total)), weight=(total))) +
   coord_flip() + 
   theme_economist() +
   labs(
     title="CBAC Spending Inceased in Between Elections",
     y = "Amount of Money Spent Per Year",
     x = "Spending Year",
     caption = "Maryland Lobbying Registrations")
```

### Question 3

**Question text**: How did the number of employers and lobbyists registered for "energy" change from 2018 to 2022?

Starting with the basic analysis of lobbyist registrations categorized under the registration matter "energy", we want to explore lobbying efforts in the energy sector. Follow-up questions include: Which firms had the most employers in our time frame? How much money did the top 3 firms make from employers in the energy sector (in 2022)? Who are their employers? Are they part of fossil fuels, clean energy or others?

Limitations: Although the general dataset on registrations goes back to 2016, datasets including the registration matter only go back to 2019. The general dataset also only goes through the end of 2021 while the energy dataset has information through November of 2022.

**Analysis summary**: As our data shows, 106 lobbyists have registered for "energy" in 2019, 132 in 2020, 123 in 2021 and 115 in 2022. One could have assumed that since the climate crisis has been big news in recent years that there would have been more lobbyists, which doesn't seem to be the case. The numbers has been reclining since 2020. Since 2020 was the first pandemic year, there could be a connection. Did employers have less money to spend on lobbying?

Interestingly, while the number of lobbyists declined, the number of employers grew before 2022. In 2019 there were 66 employers, in 2020 88, in 2021 85 and in 2022 there were 78 employers in the registration matter category "energy".

The 3 firms with the most employers registered for energy lobbying since 2019 are:
- Compass Government Relations Partners, LLC (10 employers)
- Perry White Ross & Jacobson (8 employers)
- Rifkin Weiner Livingston LLC (8 employers)

We found that Exxon Mobil Corporation appears as an employer for the first time in 2022. This is already interesting, as Exxon is a huge player in the energy sector nationwide. Further research on the MSEC website shows that Exxon has appeared in different categories on the past, "Environment" and "Others". 

But, according to his report, the lobbyist received $0. In fact, all available activity reports for Exxon indicate that the company has never spent a single dollar on lobbying in Maryland. 

According to Exxon's own 2020 lobbying report, the corporation spent the most in Alaska ($183,000) for "oil and gas issues, policies, and tax legislation". In Texas, Exxon lobbied around alcoholic beverage regulation, education, insurance, labor and malpractice-health care providers. These topics are not directly related to Exxon's business. It could be interesting to look into what Exxon is trying to achieve in Texas and how its lobbying efforts on these issues relate to its business model.

Lobbying report: https://corporate.exxonmobil.com/-/media/Global/Files/policy/lobbying/ExxonMobil_2020-Lobbying-Report.pdf

Further research on the MSEC website shows that Exxon has appeared in different categories on the past, "Environment" and "Others".

```{r}
## The dataset has been cleaned in OpenRefine. We clean the names and split the date column, so we have a start and an end date for the registration.
energy_registrations <- read_csv("data/energy_dataset/clean_energy-registrations.csv") %>% clean_names()
energy_registrations[c('start', 'end')] <- str_split_fixed(energy_registrations$registration_period, '-', 2)
```

```{r}
## Then we transform our new colums from characters to dates and create a floor date to make grouping by year possible.
energy_registrations <- energy_registrations %>% 
  mutate(new_start = as.Date(start, format = "%m/%d/%y")) %>% 
  mutate(new_end = as.Date(end, format = "%m/%d/%y")) %>% 
  mutate(start_year = floor_date(new_start, "year"))
```

```{r}
## To make sure we don't count some lobbyists twice in a year (they do sometimes register multiple times for short periods in a year), we use n_distinct.
year_energy_registrations <- energy_registrations %>% 
  group_by(start_year) %>%
  summarise(count = n_distinct(clean_lobbyist_registrant))
year_energy_registrations
```

```{r}
## We repeat the code, only this time we are looking for the number of employers lobbyists registered with.
year_energy_employers <- energy_registrations %>% 
  group_by(start_year) %>%
  summarise(count = n_distinct(clean_employer))
year_energy_employers
```

```{r}
## We want to know which firms had the most employers in our timeframe (2019-2022). We grouped by firms, counted employers and arranged the data.
employer_energy <- energy_registrations %>% 
  group_by(clean_organization_firm) %>% 
  summarise(count = n_distinct(clean_employer)) %>% 
  arrange(desc(count))
employer_energy
```
```{r}
## Let's see who those firms work for
top_three <- energy_registrations %>% 
  filter(clean_organization_firm == c("Compass Government Relations Partners, LLC", "Perry White Ross & Jacobson", "Rifkin Weiner Livingston LLC")) %>% 
  select(clean_organization_firm, clean_employer) %>% 
  arrange(clean_organization_firm)
top_three
```
```{r}
## Now we want find out who has registered for the first time in 2022.
new_registrants <- energy_registrations %>% 
  group_by(clean_lobbyist_registrant) %>% 
  summarise(first = min(start_year)) %>% 
  filter(first == "2022-01-01")
new_registrants
## And who these new registrants work for
new_registrants_employers <- energy_registrations %>% 
  group_by(clean_lobbyist_registrant, clean_employer) %>% 
  summarise(first = min(start_year)) %>% 
  filter(first == "2022-01-01")
new_registrants_employers
## Then we want to know which employers registered for "energy" for the first time in 2022
new_employers <- energy_registrations %>% 
  group_by(clean_employer) %>% 
  summarise(first = min(start_year)) %>% 
  filter(first == "2022-01-01")
new_employers
```

### Question 4

**Question text**: How did the number of Criminal Law Procedure lobbyists registered change after the September 9th youth curfew in Prince George's County? For criminal law procedure lobbyists, are there new people among the recent registrants who haven't registered before? Do any of the recently registered criminal procedure lobbyists have the same employers? 

**Analysis summary**: The number of lobbyists that registered on or after September, 01 2021 through August 2022 for Criminal Law Procedure was 25 (24 were from September 2021 to February 2022). From September, 01 2022 through November 2022, there were 19 registered lobbyists for Criminal Law Procedure. Now that we know how the number of lobbyists has changed, it would be interesting to conduct more research and see if any of them have a connection to PG County at all.

I went to the Count Sept22 to Nov data set since it already counted the distinct number of registrants from Sep. 2022 to Nov 22. I took the 19 names from that list, filtered them from the initial criminal law procedure data set, and created a new dataframe. Afterwards, I looked through the new CrimeRecentReg data set to see if there were any registrants that did not repeat.

The most recent criminal law procedure lobbyists registrants that were not previously registered in this category before were Carrington, J. Darrell, Elliott, Robyn S, Hessler, Therese Marie, Lombardi, Joyce, McDonough, Caitlin E, Zeimer, Matan.

Some of the recently lobbyists also had the same employer. Both Yanet and Lorenzo had the ACLU of Maryland as their employer. Lisa, Dorthy J, Sean R, Catlin E, and Thomas R had the House of Ruth Maryland as their employer. The only person whose employer was the Jews United for Justice was Matan. People like Kanye West and Kyrie Irving have made the news lately due to Antisemitic posts and remarks that they have made. I think it would be interesting to see if there will or has been an increase in the employers hiring lobbyists (from all lobbyists data) to advocate for the rights of the Jewish community.

Criminal Law Procedure data set

```{r}
Criminal_Prod<- read_csv("data/CriminalLProd.csv") %>% clean_names()
Criminal_Prod[c('start', 'end')] <- str_split_fixed(Criminal_Prod$registration_period, '-', 2)
NewCriminal_Prod <- Criminal_Prod %>% 
mutate(new_start = as.Date(start, format = "%m/%d/%y")) %>% 
  mutate(new_end = as.Date(end, format = "%m/%d/%y")) %>% 
  mutate(start_year = floor_date(new_start, "year")) %>% 
  group_by(lobbyist_registrant)
NewCriminal_Prod
```
Compare: March 2022 through August 2022 to September 2021 to February 2022

```{r}
SepttoFeb <- NewCriminal_Prod %>% 
filter(new_start == "2021-11-08" | new_start == "2021-11-01" |
new_start == "2022-02-08" | new_start == "2022-01-01" | new_start == "2021-12-06" | new_start == "2022-01-12" | new_start == "2022-01-24")
  
March22toAug <- NewCriminal_Prod %>% 
  filter(new_start == "2022-05-18" | new_start == "2022-08-11")
Sept22toNov <- NewCriminal_Prod %>% 
  filter(new_start == "2022-11-01" | new_start == "2022-11-07" | new_start == "2022-11-11" | new_start == "2022-11-02")
```

```{r}
CountSepttoFeb <- SepttoFeb %>% 
  group_by(lobbyist_registrant) %>% 
  summarise(count= n_distinct(lobbyist_registrant))
CountMarch22toAug <- March22toAug %>% 
  group_by(lobbyist_registrant) %>% 
  summarise(count= n_distinct(lobbyist_registrant))
CountSept22toNov <- Sept22toNov %>% 
  group_by(lobbyist_registrant) %>% 
  summarise(count= n_distinct(lobbyist_registrant))
```

```{r}
## To see if if any recent Criminal Law Procedure Lobbyists were previously registered as Criminal Law Procedure Lobbyists: 
CrimRecentReg <- NewCriminal_Prod %>% 
  filter(lobbyist_registrant == "Amanuel, Yanet" | lobbyist_registrant == "Bellamy, Lorenzo" | lobbyist_registrant == "Carrington, J. Darrell"|lobbyist_registrant == "Ciekot, Ann T." |lobbyist_registrant == "Egan, Ashley" | lobbyist_registrant == "Elliott, Robyn S" | lobbyist_registrant == "Hessler, Therese Marie" | lobbyist_registrant == "Jones, Lisa" | lobbyist_registrant == "Kraska, Martin" | lobbyist_registrant == "Lennig, Dorothy J" | lobbyist_registrant == "Lombardi, Joyce" | lobbyist_registrant == "Malone, Sean R" | lobbyist_registrant == "Manno, Roger" | lobbyist_registrant == "Mansfield, Andrea" | lobbyist_registrant == "McDonough, Caitlin E" | lobbyist_registrant == "Shapiro, Melanie" | lobbyist_registrant == "Tompsett Jr., Thomas R." | lobbyist_registrant == "Weisel, Meredith R." | lobbyist_registrant == "Zeimer, Matan")
```


### Question 5 

**Question text**: What lobbyist/registrant represents the most employers in our data set?

**Analysis summary**: The lobbyist that represents the most employers in our data set was Timothy A. Perry. He's been representing employers for 12 years as a member of Perry White Ross Jacobson LLC and served as Chief of Staff to Senate President Thomas V. Mike Miller, Jr., the longest serving State Senate President in United States history. He lives in Baltimore County, Maryland, and has represented 126 distinct employers in our joined data set. He's represented employers such as Stevenson University and AT&T over four of our data cycles. According to a profile written about him in September 2022, he specializes in government relations and serves as a board member of First Fruits Farm, a charitable nonprofit organization that specializes in food services. (https://marquistoplawyers.com/2022/09/16/timothy-perry/)

Firms with most employers: 
- Compass Government Relations Partners, LLC
- Perry White Ross & Jacobson
- Rifkin Weiner Livingston LLC
Research on website

In the below block, I counted the number of total employers that each lobbyist represents, then arranged in descending order to see who has the most total employers in the cleaned data set.
```{r}
most_employers <- clean_total_registrations %>% 
  group_by(lobbyist_registrant) %>%
  count() %>% 
  arrange(desc(n))
most_employers
```

Here, I filtered for Timothy A. Perry's different employers across each of the data sets to see the different employers he represents. There, I realized that he's represented a lot of the employers over multiple data sets, effectively skewing the above data.

```{r}
clean_total_registrations %>% 
  filter(lobbyist_registrant == "Perry, Timothy  A")
```

I used the codeblock below to see which of Perry's employers he's represented over the most datasets. For example, he's represented Anheuser-Busch Companies four different times, which makes the 333 number in the most_employers data set no longer the number I want to see. I want to see how many unique companies each lobbyist represented, and the current code accounts for all of the duplicates instead of each distinct employer.

```{r}
dupe_employers <- clean_total_registrations %>% 
   filter(lobbyist_registrant == "Perry, Timothy  A") %>% 
  count(clean_employer, source = TRUE) %>% 
  arrange(desc(n))
dupe_employers
```

By grouping by lobbyist, organization firm and employer and counting for distinct employers before ungrouping and regrouping by lobbyist, I was able to figure out how many unique employers each lobbyist represents. Perry still leads the way but with a far smaller number than the original data set after all of the duplicate employers were removed. This number better answers the original question that we took from the data, which was who represents the most UNIQUE employers in the combined data set.

```{r}
distinct_employers <- clean_total_registrations %>% 
    group_by(lobbyist_registrant, clean_organization_firm, clean_employer) %>% 
  summarise(
    count= n_distinct(lobbyist_registrant, clean_organization_firm, clean_employer)
  ) %>% 
  ungroup() %>%
  group_by(lobbyist_registrant) %>% 
  count() %>% 
arrange(desc(n))
distinct_employers
```