---
title: "lab_03"
author: "Abby Wallace"
date: "8/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse here
library(tidyverse)

# Load janitor here
library(janitor)
```

## Load data
```{r}
# Write code to load MD slice of FEC House candidate contribution data and call it md_contribs.
# The data is called lab_03.rds, and is in the data folder alongside lab_03.Rmd (this data notebook)
md_contribs <- read_rds("data/lab_03.rds")

```

## Answer questions

Q1: Which seven cities in Maryland had the largest _number_ of contributions from donors? How many did each have? To answer this question, you'll have to clean the messy "contributor_city" field.

Note: To answer this question, you'll need to see every variation of the "contributor_city" for the top seven cities, which you should identify through some initial analysis. Then find variations of them and change those variations into the most common version.

A1: The top seven cities with the most contributions are Baltimore (6726), Bethesda (5860), Sliver Spring (4659), Chevy Chase (3261), Potomac (2302), Rockville (2213), & Annapolis (2001).
```{r}
cleaned_md <- md_contribs %>%
  mutate(
    contributor_city = str_to_upper(contributor_city)) %>%
    mutate(
      contributor_city = case_when(
        contributor_city == "BALITMORE" ~ "BALTIMORE",
        contributor_city == "BETHSDA" ~ "BETHESDA",
        contributor_city == "SILVER SPRINGS" ~ "SILVER SPRING",
        contributor_city == "ANNNAPOLIS" ~ "ANNAPOLIS",
        contributor_city == "BETHESA" ~ "BETHESDA",
        contributor_city == "POTAMIC" ~ "POTOMAC",
        contributor_city == "11509 LAKE POTOMAC DRIVE" ~ "POTOMAC",
        contributor_city == "ANAPOLIS" ~ "ANNAPOLIS",
        contributor_city == "ANNAPOLID" ~ "ANNAPOLIS",
        contributor_city == "ANRETIREDPOLIS" ~ "ANNAPOLIS",
        contributor_city == "CHEVY CHASE VILLAG" ~ "CHEVY CHASE",
        contributor_city == "SILVER SPRING, MD" ~ "SILVER SPRING",
        contributor_city == "SLIVER SPRING" ~ "SILVER SPRING",
    TRUE ~ contributor_city
      ))
cleaned_md %>%
  group_by(contributor_city) %>%
summarise(
    count=n()
  ) %>%
  arrange(desc(count))  

```

Q2: Find the 5-digit ZIP code in Maryland that had the highest total amount of contributions (by dollar value). Where is that zip code and what kind of area is it? Which committee is the largest recipient of money from that zip code? How many of the top five recipients are Maryland candidates?

A2: ZIP code 20815 had the highest dollar value amount of contributions at $1804495.47. Cities in this zip code include  Chevy Chase, Bethesda, Silver Spring, Friendship Village, Somerset, & Martin's Additions. JAMIE RASKIN FOR CONGRESS is the committee reciving the most money from this ZIP code at $178088.55.	Jamie Raskin and Heather Mizeur are the only Maryland-based politicians out of the top five committees receiving contributions from this ZIP code. Susan Wild, Nancy Pelosi, and Liz Cheney are all out of state.

```{r}
# Hint
# First, create a dataframe that shows the top 5-digit ZIP code by total dollar amount

cleaned_md <- cleaned_md %>%
   mutate(contributor_zip = str_sub(contributor_zip, start=1L, end=5L))

cleaned_md %>%
group_by(contributor_zip) %>%
summarise(
    total_amount = sum(contribution_receipt_amount)
  ) %>%
  arrange(desc(total_amount))  

```
```{r}
# Hint
# Second, use the top ZIP code in the dataframe you just created and create a new dataframe that shows the committee that got the most money from that ZIP code.
filtered_zip <- cleaned_md %>% filter(contributor_zip == "20815")
filtered_zip %>%
group_by(committee_name) %>%
summarise(
    total_amount = sum(contribution_receipt_amount)
  ) %>%
  arrange(desc(total_amount))  

```

Q3: Let's look at top individual donors to House campaigns. Write code to generate a list of top donors using name and city, adding up the total amount of contributions. One of the top few donors has a very strong connection to the U.S. House - who is that and what does he do? How does his donation pattern differ from the second-largest donor, and what can you say about that second-largest donor?

Feel free to add any interesting "color" (little interesting details) you find in your research.  

A3: The donor who contributed the highest dollar amount of contributions is David Trone - who is the Democratic representative of Maryland's 6th Congressional district. Trone's 35 donations seem to be going to other democratic colleagues around the country and feature pretty similar donation amounts/ranges. Juanita Mason, however has over 3072 contributions seem to have more sporatic amounts of money with the top committees being out of state Republican candidates.

```{r}
# Hint
# First, create a dataframe of top individual contributors
cleaned_md %>%
group_by(contributor_name, contributor_city) %>%
summarise(
    total_amount = sum(contribution_receipt_amount),
    count_contribs = n()
  ) %>%
  arrange(desc(total_amount))  

```

```{r}
#Hint
# Second, use information from that dataframe to examine individual donations from the original md_contribs dataframe.

filtered_contributor <- cleaned_md %>% filter(contributor_name == "TRONE, DAVID")
filtered_contributor %>%
group_by(committee_name) %>%
summarise(
    total_amount = sum(contribution_receipt_amount)
  ) %>%
  arrange(desc(total_amount)) 
```

```{r}
#Hint
# Second, use information from that dataframe to examine individual donations from the original md_contribs dataframe.

filtered_contributor <- cleaned_md %>% filter(contributor_name == "MASON, JUANITA")
filtered_contributor %>%
group_by(committee_name) %>%
summarise(
    total_amount = sum(contribution_receipt_amount)
  ) %>%
  arrange(desc(total_amount)) 
```
