---
title: "lab_04"
author: "Abby Wallace"
date: "10/07/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings
```{r}
# Load the libraries you will need to read in and clean up the data, knowing that you can go back and add a library and re-run this block after you get a look at the dataframes.
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse here
library(tidyverse)

# Load janitor here
library(janitor)

library(lubridate)
```

## Getting Started

This week we'll be combining and joining data to make it more useful and to ask some questions about it. We'll use Maryland 2022 primary precinct election results to do this, and our main question will be this: Which precincts had the most votes cast at them statewide, counting both the Republican and Democratic primaries? To answer this question, we'll need to first combine the two parties' primary precinct results into a single dataframe and then join it to data that has more information about each voting location.

First, load the two precinct results CSV files in the `data` folder, one each for the Democratic and Republican primaries, and then combine them into a single dataframe using the method we used in the pre_lab. You'll need to clean up the column names to make the data easier to work with - make sure you have loaded the library to do that.

```{r}
dem_results <- read_csv("data/md_primary_precinct_dem.csv")
```
```{r}
rep_results <- read_csv("data/md_primary_precinct_rep.csv")
```

Next, let's read in the precinct voting location data and make sure you clean up the column names:

```{r}
precinct_locations <- read_csv("data/precinct_voting_locations.csv")
```

These two dataframes have a pair of columns in common - not necessarily with the same column names. Join the two dataframes on that unique combination that describes a precinct and the jurisdiction that contains it:

```{r}
combined_results <- bind_rows(list(dem_results, rep_results))

View(combined_results)
```

## Answer questions

Q1. Using the newly-created dataframe and isolating just the results for Wes Moore and Aruna Miller, write code that calculates how many voting locations each county has, showing the county with the most voting locations first. There's a catch: multiple precincts can share the same voting location, so you'll need to calculate the distinct number of locations (Google is your friend).

Does the county with the most voting locations have the largest population in the state (ask yourself: where have you seen county population figures recently)?

A1. No, the most populated county is Montgomery County which has the second most polling places with 225. According to the code from the table, Prince George's County has the most polling places with 240.

```{r}
cleaned_combined <- combined_results  %>%
  clean_names()

filtered_results <- cleaned_combined %>% filter(candidate_name == "Wes Moore and Aruna Miller")
filtered_results
```

```{r}
changed_precincts <- precinct_locations %>%
  mutate(PRECINCT_CODE = as.numeric(PRECINCT_CODE))
changed_precincts
```

```{r}
joined_results <- filtered_results %>% left_join(changed_precincts, by=c("county"="COUNTY_CODE", "election_district_precinct"="PRECINCT_CODE"))

```

```{r}
joined_results %>% 
  group_by(county_name) %>%
  summarize(count = n_distinct(LOCATION_NAME)) %>%
  arrange(desc(count))
  
```

Q2. Write code to produce a `votes` column which is the result of adding together `early_votes`, `election_night_votes`, `provisional_votes` and the mail-in votes, overwriting your dataframe so that it now has that `total_votes` column going forward. Then, separately write code to show each county and precinct and calculate the total votes in the Governor's race - bearing in mind that each precinct will have multiple rows in the dataframe, but you only want to see one row for each precinct. What is the name and voting location of the precinct that saw the most votes in the governor's race? 

A2. Dr. Henry A. Wise, Jr. High School was the location and Prince George's was the county for Precinct 15005. This precinct had the most votes for the governor's race at 2396.

```{r}
q2_votes <- cleaned_combined %>%
  summarise(
    total_early = sum(early_votes),
    total_election_night = sum(election_night_votes),
    total_provisional = sum(provisional_votes),
    total_mail_in = sum(mail_in_ballot_1_votes + mail_in_ballot_2_votes),
    votes = sum(total_provisional + total_election_night + total_early + total_mail_in))
complete_votes <- q2_votes %>%
    rename(total_votes = votes)
  
complete_votes
```
```{r}
joined_gov <- cleaned_combined %>% left_join(changed_precincts, by=c("county"="COUNTY_CODE", "election_district_precinct"="PRECINCT_CODE")) %>%
  group_by(county_name, election_district_precinct, LOCATION_NAME) %>% 
  filter(office_name == "Governor / Lt. Governor") %>% 
  summarise(total_votes = sum(early_votes, provisional_votes, mail_in_ballot_1_votes, mail_in_ballot_2_votes, election_night_votes)) %>% 
  arrange(desc(total_votes))

joined_gov
```

Q3. Let's look at the average number of votes in each party's primary race for governor. Take the code you wrote for the last part of Q2, save the results to a new variable and then find the precincts with 0 total votes for governor. How many precincts are there, and which county has the most? What else would you need to know about these precincts to know whether this might be worth a story?

Finally, write code to calculate the average and median number of total votes across all precincts statewide (so you want to see just two numbers - the average and median). To do this, you will need to use the `ungroup()` function on your new dataframe _before_ you do the math - https://www.rdocumentation.org/packages/multiplyr/versions/0.1.1/topics/ungroup. There's no state column here, so you aren't grouping by anything, but can still use summarize to do the math.

Take a look at the the Annotated Code of Maryland's election laws (https://mgaleg.maryland.gov/2023RS/Statute_Web/gel/gel.pdf) - can you find a limit on the number of voters assigned to a precinct? How might your answer impact wait times to vote?

A3. If you look at the joined_gov full table from the last question and sort by total_votes, you can see 12 precincts with 0 total votes for governor. Prince George's County has the most precincts with 0 votes for governor with 4. I would want to know voter demographics for these precincts. What races did those voters partake in? Why didn't they feel the need to vote for governor? How many voters came to each of these precincts? Also, I did not see a limit to the number of voters assigned to a precinct. The mean for total votes across the state was 465.9 and the median total was 418. I feel like that is when the voting locations could get overfilled and have such long lines - too many people in too small of a precinct.

```{r}
joined_gov %>%
  ungroup() %>%
  summarise(
    mean_total=(mean(total_votes)),
    median_total=(median(total_votes))
  )

```

Q4. Choose your own adventure: I want you to come up with a question you could ask of one of the dataframes we've created for this lab. It could involve candidates, or counties or types of voting, but it cannot be a variation of a previous question (so, for example, you can't ask, "What's the average number of votes per precinct in the U.S. Senate primary?"). First write the question and then try to answer it by writing code. If you're not sure of the exact code, try googling (or ask me or Ramsay for advice) but, at a minimum, describe what you think you'd have to do to answer it and what the results would look like. That can help.
A4: Put your question here. What candidate and precinct received the highest number of votes in Baltimore County for Attorney General? Who had the lowest total number of votes from that same precinct? Answer: Anthony G. Brown had the highest total votes in a precinct in Baltimore County with 897 votes from precinct 2010. In precinct 2010, the candidate with the lowest number of total votes was Jim Shalleck with 17 votes.

```{r}
joined_ag <- cleaned_combined %>% left_join(changed_precincts, by=c("county"="COUNTY_CODE", "election_district_precinct"="PRECINCT_CODE")) %>%
  group_by(election_district_precinct, candidate_name) %>% 
  filter(
    (office_name == "Attorney General"),
    (county_name == "Baltimore County")) %>% 
  summarise(total_votes = sum(early_votes, provisional_votes, mail_in_ballot_1_votes, mail_in_ballot_2_votes, election_night_votes)) %>% 
  arrange(desc(total_votes))
joined_ag
```
